Can Auxiliary Indicators Improve COVID-19 Forecasting and Hotspot Prediction?

### 2021-10-31 forecast/paper/supplement-text.tex
Rebuilt with edits from me.

\chngcli~(and, to a lesser extent, the other claims-based signals) is the most
affected by this distinction, reflecting the latency in claims-based reporting.
This underscores the importance of efforts to provide ``nowcasts'' for claims
signals (which corresponds to a 0-ahead forecast of what the claims signal's
value will be once all data has been collected). Looking at the \chngcli~and
\dv~curves in Figure \ref{fig:fcast-finalized}, we can see that they perform
very similarly when trained on the finalized data.  This is reassuring because
they are, in principle, measuring the same thing (namely, the percentage of
outpatient visits that are primarily about COVID-related symptoms), but based on
data from different providers.  The substantial difference in their curves in
Figure 3 of the main paper must therefore reflect their having very different
backfill profiles.

@@

Our goal in this section is to compare the AR model and indicator models to
those in the Hub, in terms of forecast errors aggregated over the same forecast
tasks, to give a clear sense for how robust and effective the models we choose
to investigate in the paper are relative to those in common operational use.
This was prompted by a question from an anonymous reviewer of this paper, who
asked why we chose to build our analysis of indicator utility around the AR
model in the first place, and why we did not build it around others (say, the
SIR model or more complex mechanistic models of disease transmission) that have 
occupied more of the spotlight over the course of the pandemic.  The analysis
presented here corroborates the claim that, while simple, the AR model, properly
trained---using a quantile loss to directly estimate multiple conditional
quantiles, a trailing training window of 21 days, pooling across all locations
jointly, and fitting to case rates rather than counts (as we do in all our
models in the main paper)---can be robust and effective, performing
competitively many of the top models from COVID-19 Forecast Hub, including the
Hub ensemble model.

@@

Therefore, to make a comparison to models in the Hub as direct as possible, we
retrained our models over the same forecast period as in the main paper, and
with the same general setup entirely, except at the state rather than HRR level.
We then rescaled them post hoc to account for the different temporal resolution
and the rate versus count scale (first and third points in the above list).  The
results are given in Figure \ref{fig:compare-to-hub}.  The evaluation was
carried out exactly as in the main paper, and the figure displays both mean WIS
and geometric mean WIS, as a function of ahead, relative to the baseline model.
Furthermore, to account for missingness (not all teams submitted forecasts to
the Hub for all locations and ahead values for the entire period), we first
dropped any forecaster than submitted for less than 6 weeks, and then restricted
the aggregation metrics (mean or geometric mean) to errors from commonly
available forecast tasks.

By either metric, mean or geometric mean WIS relative to baseline, we can see  
in Figure \ref{fig:compare-to-hub} that the AR model examined in this paper is 
competitive with top models in the Hub, even outperforming the Hub ensemble
model for smaller ahead values.  The same general conclusion can be drawn for
the indicators models as well.  However, interestingly, a close inspection
reveals that the AR model here is for the most part in the ``middle of the
pack'' when compared to the indicator models, and only the Google-AA model
offers clear improvement over AR for all aheads.  This is likely due to the fact
that at the state level, the signal-to-noise ratio (SNR) is generally higher,
and AR model provides a higher standard (on which to expect improvement using an 
auxiliary indicator), since it is able to extract a clearer signal from past
lags of case rates.  At the HRR level, with lower SNR, using indicators as
simple additional linear features in the AR model probably leads to a variance
reduction that is enough to boost accuracy, but at the state level, perhaps more
sophisticated modeling techniques are needed to extract comparable value from
some of the indicators.

>>>>

\chngcli~(and, to a lesser extent, the other claims-based signals) is the most
affected by this distinction, reflecting the latency in claims-based reporting.
This highlights the importance of efforts to provide ``nowcasts'' for claims
signals (which corresponds to a 0-ahead forecast of what the claims signal's
value will be once all data has been collected). Looking at the \chngcli~and
\dv~curves in Figure \ref{fig:fcast-finalized}, we can see that they perform
very similarly when trained on the finalized data.  This is reassuring because
they are, in principle, measuring the same thing (namely, the percentage of
outpatient visits that are primarily about COVID-related symptoms), but based on
data from different providers.  The substantial difference in their curves in
Figure 3 of the main paper must, therefore, reflect their having very different
backfill profiles.

@@

Our goal in this section is to compare the AR model and indicator models to
those in the Hub, in terms of forecast errors aggregated over the same forecast
tasks, to give a clear sense for how robust and effective the models we choose
to investigate in the paper are relative to those in common operational use.
This was prompted by a question from an anonymous reviewer of this paper, who
asked why we chose to build our analysis of indicator utility around the AR
model in the first place, and why we did not build it around others (say, the
SIR model or more complex mechanistic models of disease transmission) that have 
occupied more of the spotlight over the course of the pandemic.  The analysis
presented here corroborates the claim that, while simple, the AR model, properly
trained---using a quantile loss to directly estimate multiple conditional
quantiles, a trailing training window of 21 days, pooling across all locations
jointly, and fitting to case rates rather than counts (as we do in all our
models in the main paper)---can be robust and effective, performing
competitively to the top models submitted to the COVID-19 Forecast Hub,
including the Hub's ensemble model.

@@

Therefore, to make the comparison to models in the Hub as direct as possible, we
retrained our models over the same forecast period as in the main paper, and
with the same general setup entirely, except at the state rather than HRR level.
We then rescaled them post hoc to account for the different temporal resolution
and the rate-versus-count scaling (first and third points in the above list).
The 
results are given in Figure \ref{fig:compare-to-hub}.  The evaluation was
carried out exactly as in the main paper, and the figure displays both mean WIS
and geometric mean WIS, as a function of ahead, relative to the baseline model.
Furthermore, to account for missingness (not all teams submitted forecasts to
the Hub for all locations and ahead values for the entire period), we first
dropped any forecaster than submitted for less than 6 weeks, and then restricted
the aggregation metrics (mean or geometric mean) to errors from commonly
available forecast tasks.

By either metric, mean or geometric mean WIS relative to baseline, we can see  
in Figure \ref{fig:compare-to-hub} that the AR model examined in this paper is 
competitive with top models in the Hub, even outperforming the Hub ensemble
model for smaller ahead values.  The same general conclusion can be drawn for
the indicator-assisted models as well.  However, interestingly, a close inspection
reveals that the AR model here is for the most part in the ``middle of the
pack'' when compared to the indicator models, and only the Google-AA model
offers clear improvement over AR for all aheads.  This is likely due to the fact
that at the state level, the signal-to-noise ratio (SNR) is generally higher,
and AR model provides a higher standard (on which to expect improvement using an 
auxiliary indicator), since it is able to extract a clearer signal from past
lags of case rates.  At the HRR level, with lower SNR, using indicators as
simple additional linear features in the AR model probably leads to a variance
reduction that is enough to boost accuracy, but at the state level, perhaps more
sophisticated modeling techniques are needed to extract comparable value from
some of the indicators.

### 2021-10-31 forecast/paper/supplement.Rmd
Rebuilt with edits from me.

```{r compare-to-hub, fig.cap="Forecast performance for AR and indicators models, each retrained at the state level, compared to models submitted to the COVID-19 Forecast Hub over the same period. The thin grey lines individual models from the Hub; the blue line is the Hub ensemble model. (Note that, to align prediction dates as best as possible, we look at the AR and indicator model forecasts for 5, 12, 19, and 26 days ahead; this roughly corresponds to 1, 2, 3, and 4 weeks ahead, respectively, since in the Hub, models typically submit forecasts on a Tuesday for the epiweeks aligned to end on each of the following 4 Saturdays.)"}
knitr::include_graphics("fig/compare-states-to-hub.pdf")
```

>>>>

```{r compare-to-hub, fig.cap="Forecast performance for AR and indicator models, each retrained at the state level, compared to models submitted to the COVID-19 Forecast Hub over the same period. The thin grey lines are individual models from the Hub; the blue line is the Hub ensemble model. (To align prediction dates as best as possible, we look at the AR and indicator model forecasts for 5, 12, 19, and 26 days ahead; this roughly corresponds to 1, 2, 3, and 4 weeks ahead, respectively, since in the Hub, models typically submit forecasts on a Tuesday for the epiweeks aligned to end on each of the following 4 Saturdays.)"}
knitr::include_graphics("fig/compare-states-to-hub.pdf")
```

### 2021-10-31 forecast/paper/supplement.tex
Rebuilt with edits from me.

\caption{Forecast performance for AR and indicators models, each retrained at the state level, compared to models submitted to the COVID-19 Forecast Hub over the same period. The thin grey lines individual models from the Hub; the blue line is the Hub ensemble model. (Note that, to align prediction dates as best as possible, we look at the AR and indicator model forecasts for 5, 12, 19, and 26 days ahead; this roughly corresponds to 1, 2, 3, and 4 weeks ahead, respectively, since in the Hub, models typically submit forecasts on a Tuesday for the epiweeks aligned to end on each of the following 4 Saturdays.)}\label{fig:compare-to-hub}
\end{figure}

>>>>

\caption{Forecast performance for AR and indicator models, each retrained at the state level, compared to models submitted to the COVID-19 Forecast Hub over the same period. The thin grey lines are individual models from the Hub; the blue line is the Hub ensemble model. (To align prediction dates as best as possible, we look at the AR and indicator model forecasts for 5, 12, 19, and 26 days ahead; this roughly corresponds to 1, 2, 3, and 4 weeks ahead, respectively, since in the Hub, models typically submit forecasts on a Tuesday for the epiweeks aligned to end on each of the following 4 Saturdays.)}\label{fig:compare-to-hub}
\end{figure}
